\section{Conclusions}
\paragraph{Issues}
The major issue that came up during the course of this work was coming up with an acceptable consistency model that can be implemented. We made a few simplifying assumptions on communication reliability for the semantics, while building in support that allows modeling varying degrees of unreliability in communication. Our experiments also only involved observing limited executions of the programs and pattern matching for bad configurations. We were unable to come up with a reasonable verification procedure from the \K implementation, and that is a problem that has currently been pushed to the future work section. 
We were also unable to come up with a formal semantic definition for the DoReachAvoid abstraction, which is a major feature provided by the \StarL framework. One of the problems with coming up with a formal semantic definition for \verb|doReachAvoid| is the way we model time. Since we cannot simulate continuous time evolution, we cannot reliably capture behaviors in which sensors detect collision and robots perform a back-off maneuver. One of the solutions we are currently exploring is using a different tool to perform this simulation and feed the resulting configuration back to the \K framework. 

\paragraph{Future Work}
We are currently extending the formal semantics of the language to include the implementation of neighborhood based address information as mentioned earlier, and a publish-subscribe model for shared memory. We also plan to implement a executable semantics for reliable motion simulation using an off-the-shelf tool like C2E2\RG{citation}; as well as add support for different families of differential equations of motion. 
We only observed limited runs of the programs and pattern matched for `bad' or unsafe configurations in them, but we plan to employ and extend the verification tools \K provides; for instance symbolic execution to verify the correctness of these applications. We are currently only looking at invariants, and future work can involve exploring progress guarantees as well. 