\subsection{Overview}
\label{sect:Overview}
\rolang allows users to write applications that will run on a distributed system of agents. The user writes a program as though for a single agent, and all agents execute the same code. A \rolang program is a collection of \emph{declarations} and \emph{events}. Shared variables are used for communication between agents. They are declared in an \verb|MW| block. \verb|MW| stands for \emph{multi-writer}, implying that all agents can read from, and write to the variables declared in this block. We provide another type of shared variables called \emph{shared single-writer} variables, which all agents can read from, but only one agent can write to. These variables are parameterized by the \verb|agent index|, an integer which is a unique identifier for each agent in the system, and they are declared in the \verb|SW| block. Local variables are declared in \verb|Loc|al declaration blocks.  \newline

As mentioned earlier, \rolang uses a \emph{precondition-effect} style of programming. These precondition and effect statements form events. Each application consists of a special \verb|Init| event, and an \verb|EventBlock|. The \verb|Init| event occurs at the beginning when the application starts executing, and it contains all statements which need to be executed only once; for instance, initializing a shared array. After the \verb|Init| event is executed, the \verb|EventBlock| starts executing. It contains a list of events that define the behavior of the application. The preconditions of each of the events inside the \verb|EventBlock| are evaluated in order of appearance, and the effect is executed if the precondition becomes true. The \verb|EventBlock| can be seen as a (potentially) infinite while-loop. 

We also provide an abstraction to manage the physical control of the agents, called \verb|doReach|, which takes as input a \emph{target} to reach, and a list of (predetermined) \emph{obstacles} which need to be avoided. We do not need to specify the format of the obstacles, as different implementations of this \verb|doReach| can have different specifications, but the target in general has the same type as the time varying variables of the system. \verb|doReach| communicates with the program using two flags, \emph{doReach_done}, and \emph{doReach_fail}. If the target is reached, the \emph{doReach_done} is set to true, and the \emph{doReach_fail} is set to true when the agent does not seem to have reached the target.

The next section presents the formal syntax, and an example to illustrate the structure of a general application. 

\input{syntax.tex}